---
title: "example"
output: html_document
date: "2025-06-16"
---

Load example data.
```{r}
# library("MetabAnnotate")
data("exampleData")
```

Estimate the number of latent factors, then transform the input data into the format required by MetabAnnotate:

We recommend use BEMA for raw data, and dBEMA for summary-level data.
The BEMA function is part of the package and was downloaded from the GitHub repository maintained by ZhengTracyKe, available at: https://github.com/ZhengTracyKe/BEMA.
```{r}
Y.copy = na.omit(Y)#remove rows including NAs in Y
K = BEMA(data = t(Y.copy),n = ncol(Y.copy),p = nrow(Y.copy),alpha = 0.1)

mtb_data = preProcess(IDs = IDs,Y = Y,K = K,impute = T,imputation.methods = "Mean",COMD.column = "CHEMICAL_ID")
```

The dBEMA function is part of the package and was obtained from the GitHub repository maintained by Weiqiong Huang, available at: https://github.com/WeiqiongHuang/HiGSS.
```{r}
est.K = dBEMA(Stand.B = B,N = N,alpha = 0.1,n.avg = 5)# Only 5 simulation to estimate the bulk eigenvalue distribution for time efficiency.
K = est.K$nFactors

mtb_data = preProcess(IDs = IDs,B = B,N = N,K = K,COMD.column = "CHEMICAL_ID")
```

Implement the annotation function by randomly selecting 20 metabolites for calibration. Generate 20 Gibbs samples, discarding the first 10 iterations as burn-in. It is recommended to run 100 iterations and 50 burn-in for real annotation tasks (Iter = 100, burnin = 50).
```{r}
test = metabAnnotate(mtb_data = mtb_data,size.clbr = 20,Iter = 20,burnin = 10,new.pathway = F,seed = 1)
```

Check the annotation results.
```{r}
head(test$annt.result)
```

